Report on a Safety Checker for Petri nets

### 2014-02-01 paper/workspace/main.tex
bla
Changed example in paper draft to subnet of Lamport's algorithm

\includegraphics[scale=0.75]{lamport-code}

\begin{verbatim}
* Property: Thread 1 and Thread 2 are never in the critical section at the same time.

@@

\begin{center}
  \includegraphics[width=12cm]{lamport}
\end{center}

\begin{verbatim}
* Property P: cs1 + cs2 <= 1
\end{verbatim}

\newpage
\begin{verbatim}
* Method Safety:

@@

\begin{verbatim}
* Property of state constraints C: If C U {\neg P} is unsat then N |= P.

* Example:
  - State constraints C:

@@

    id1        = 1 + u1 - u2
    ^            ^   ^    ^
    |            |   |    |
    |            |   |    |
    |            |   |    |
    |            |   |   # of tokens taken from nc
    |            |   |    
    |            |  # of tokens given to nc
    |            |
    |          initial number of tokens in nc
    |
    number of tokens in nc

    cs1        = 0 - u1      + u3
    idle1      = 1 + u1 - u2
    req1       = 0      + u2 - u3
    nid1       = 0 - u1 + u2
    id2        = 1                     + v2           + v5 - v6
    cs2        = 0                               + v4 - v5
    idle2      = 1                          + v3 + v5      - v6
    req2       = 0                - v1           - v4      + v6
    after_you2 = 0                + v1 - v2
    await2     = 0                     + v2 - v3
\end{verbatim}

\begin{verbatim}
    Non-negativity conditions:

    id1        >= 0
    cs1        >= 0
    idle1      >= 0
    req1       >= 0
    nid1       >= 0
    id2        >= 0
    cs2        >= 0
    idle2      >= 0
    req2       >= 0
    after_you2 >= 0
    await2     >= 0
    u1         >= 0
    u2         >= 0
    u3         >= 0
    v1         >= 0
    v2         >= 0
    v3         >= 0
    v4         >= 0
    v5         >= 0
    v6         >= 0

@@

    cs1 + cs2  >= 2

@@

\newpage
\begin{verbatim}
* Example
  - State constraints C and \neg P are the same as in the example for method Safety.

  - Assignment A:
    id1        = 0
    cs1        = 1
    idle1      = 0
    req1       = 0
    nid1       = 1
    id2        = 0
    cs2        = 1
    idle2      = 0
    req2       = 0
    after_you2 = 0
    await2     = 0
    u1         = 0
    u2         = 1
    u3         = 1
    v1         = 0
    v2         = 0
    v3         = 0
    v4         = 1
    v5         = 0
    v6         = 1

@@

    - Trap implications:
                       v4              v3                 u2
                  ------------    --------------    --------------     
    id1        => (cs2 \/ id1) /\ (id1 \/ idle2) /\ (req1 \/ nid1)
    ^              ^      ^        ^      ^          ^       ^
    |              |      |        |      |          |       |
    |              |     id1 \in S |     idle2 \in S |      nid1 \in S
    |              |               |                 |
    |             cs2 \in S       id1 \in S         req1 \in S
    |
    id1 \in S

    cs1        => idle1 \/ id1
    idle1      => req1 \/ nid1
    req1       => cs1 \/ id2
    nid1       => (after_you2 \/ nid1) /\ (idle1 \/ id1)
    id2        => req2 /\ (cs1 \/ id2)
    cs2        => (idle2 \/ id2)
    idle2      => req2
    req2       => (cs2 \/ id1) /\ (cs2 \/ id1)
    after_you2 => id2 \/ await2
    await2     => id1 \/ idle2

    - At least one of the initially marked places belongs to S:
    id1 \/ id2 \/ idle1 \/ idle2

    - None of the marked places in A belongs to S:
    ~cs1 /\ ~cs2 /\ ~nid1
\end{verbatim}

@@

\begin{verbatim}
  - Assignment A_\theta:
    id1        = true
    cs1        = false
    idle1      = false
    req1       = true
    nid1       = false
    id2        = true
    cs2        = false
    idle2      = false
    req2       = true
    after_you2 = true
    await2     = false

  - Refinement constraint \delta:
    id1 + id2 + req1 + req2 + after_you2 >= 1
    ^     ^     ^      ^      ^
    |     |     |      |      |
    ---------------------------
      |
      S = {iq1, iq2, req1, req2, after_you2}, therefore \delta excludes
      assignment A in the next iteration

>>>>

Process 1:

bit1 := false
while true do
  bit1 := true
  a: while bit2 do skip od
  b: <critical section>
  bit1 := false
od

Process 2:

bit2 := false
while true do
  l: bit2 := true
  if bit1 then
    bit2 := false
    while bit1 do skip od
    goto l
  fi
  <critical section>
  bit2 := false
od

* Property: Process 1 and Process 2 are never in their respective critical section at the same time.

@@

\begin{center}
  \includegraphics[width=10cm]{lamport}
\end{center}

* Subnet of Petri net:

\begin{center}
\begin{tikzpicture}[node distance=2cm]
  \node[place, tokens=1, label=left:a] (a) {};
  \node[place, tokens=0, label=left:b, below of=a] (b) {};
  \node[place, tokens=0, label=right:nbit2, right of=b, xshift=-1cm, yshift=+1cm] (nbit2) {};
  \node[transition, label=left:t1, above of=a, yshift=-1cm] (t1) {}
    edge [pre, bend left, min distance=2cm, out=225, in=-45] (b)
    edge [post] (a);
  \node[transition, label=left:t2] (t2) at ($(a)!0.5!(b)$) {}
    edge [pre] (a)
    edge [pre] (nbit2)
    edge [post] (b)
    edge [post] (nbit2);
\end{tikzpicture}
\end{center}

\begin{verbatim}
* Property P: a + b <= 1
\end{verbatim}

\begin{verbatim}
* Method Safety:

@@

\begin{verbatim}
* Property of state constraints C: If C U {\neg P} is unsat then N |= P.
\end{verbatim}

\begin{verbatim}
* Example:
  - State constraints C:

@@

    a     = 1 + t1 - t2
    ^       ^   ^    ^
    |       |   |    |
    |       |   |    |
    |       |   |    |
    |       |   |   # of tokens taken from a
    |       |   |    
    |       |  # of tokens given to a
    |       |
    |      initial number of tokens in a
    |
    number of tokens in a

    b     = 0 - t1 + t2
    nbit2 = 0

Non-negativity conditions:

    a     >= 0
    b     >= 0
    nbit2 >= 0

@@

    a + b >= 2

@@

\newpage
\begin{verbatim}
* Example
  - State constraints C same as in the example for method Safety.

  - Property P:

    b = 0

    b >= 1

  - Assignment A:
    a     = 0
    b     = 1
    nbit2 = 0
    t1    = 0
    t2    = 1

@@

    - Trap implications:
               t2
             ----------
    a     => b \/ nbit2
    ^           ^      ^
    |           |      |
    |           |     nbit2 \in S
    |           |
    |          b \in S
    |
    a \in S

    nbit2 => b \/ nbit2
    b     => a

    - At least one of the initially marked places belongs to S:
    a

    - None of the marked places in A belongs to S:
    ~b

  - Assignment A_\theta:
    a     = true
    b     = false
    nbit2 = true

  - Refinement constraint \delta:
    a + b >= 1
    ^   ^
    |   |
    -----
      |
      S = {a, b}, therefore \delta excludes assignment A in the next iteration
\end{verbatim}

@@
